\documentclass[../Main.tex]{subfiles}
\setcounter{chapter}{15}

\begin{document}
\phantomsection
\chapter{L16: \texorpdfstring{$R$}{R}-module homomorphisms}
\begin{dfn}[title = {\texorpdfstring{$\Hom_R(M,N)$}{Hom}, Kernel, Image, Isomorphism}]
The s\textbf{et of $R$-module homomorphisms from $M$ to $N$} is denoted $\Hom_R(M,N)$.\\
The \textbf{kernel} of an $R$-module homomorphism $f\in \Hom_R(M,N)$ is 
\[\Ker f \coloneqq \{m\in M \mid f(m)=0\}\]
The \textbf{image} of $f\in \Hom_R(M,N)$ is
\[\Img f \coloneqq \{n\in N\mid \exists m\in M, \, f(m)=n\}\]
If $f\in\Hom_R(M,N)$ is bijective then we say $f$ is an \textbf{isomorphism of $R$-modules}. We say $M,N$ are \textbf{isomorphic} if there is an isomorphism $f\colon M\to N$ and we write $M\cong N$.
\end{dfn}
\begin{example}
	$R=\Z,M=\Z$ is a $\Z$-module.\\
	What do the $\Z$-module homomorphisms from $\Z$ to $\Z$ look like?
	\begin{align*}
	f\colon \Z \to& \Z \\
	1\mapsto& a \\
	n\mapsto& \underbrace{a+a+a+\dots+a}_{n\text{-times}}
	\end{align*}
\end{example}
\Note The doubling function
\begin{align*}
f_2\colon \Z\to& \Z \\
1\mapsto& 2
\end{align*}
is a $\Z$-module homomorphism as
\begin{align*}
f_2(m+n)&=2\rdot (m+n)=2\rdot m+2\rdot n = f_2(m)+f_2(n)\\
f_2(a\rdot m)&=2\rdot (a\rdot m) = a\rdot 2\rdot m = af_2(m)
\end{align*}
However it is \textbf{not} a ring homomorphism as
\begin{align*}
f_2(2\rdot 3) = 2\rdot 2\rdot 3 =12 \ne 24 = 4\rdot 6 = (2\rdot 2) \rdot  (3\rdot 2) = f_2(2)\rdot f_2(3)
\end{align*}
\begin{prop}[title = Kernel and Image are submodules]
	Suppose $f\in \Hom_R(M,N)$. Then the kernel $\Ker f \subset M$ and the image $\Img f \subset N$ are $R$-submodules.
\end{prop}
\begin{proof}
	First we prove the claim on the kernel. If $a,b \in \Ker f$ and $r\in R$ then
	\begin{itemize}
	\item $f(0)=0\implies 0 \in \Ker f$
	\item $f(a+b) = f(a)+f(b) = 0 +0=0 \implies a+b \in \Ker f$
	\item$f(r\rdot a) = r\rdot f(a) = r\rdot 0=0 \implies r\rdot a \in \Ker f$
	\item$0=f(0)=f(a+(-a)) = f(a) + f(-a) = 0 +f(-a) = f(-a) \implies -a\in \Ker f$
	\end{itemize}
	hence, $\Ker f \subset M$ is a submodule.\\
	If $a,b\in \Img f, r\in R$ say $a=f(a'), b=f(b'), a',b' \in M$
	\begin{itemize}
		\item $f(0)=0\implies 0\in \Img f$
		\item $a+b=f(a')+f(b')=f(a'+b')\implies a+b\in \Img f$
		\item $ra=r\rdot f(a')=f(r\rdot a') \implies r\rdot a \in \Img f$
		\item $-a=-f(a')=f(-a') \implies -a\in \Img f$
	\end{itemize}
	Hence, $\Img f$ is a submodule.
\end{proof}
\begin{dfn}[title = coset]
	If $N\subset M$ is an $R$-submodule and $m\in M$, then the \textbf{$N$ coset} of $m$ is
	\[m+N \coloneqq \{m+n\mid n\in N\}\]
\end{dfn}
\Exr We can define an equivalence relation on $M$ by $m\sim m'$ if and only if $m+N = m'+N$ as sets.
\begin{dfn}[title = Quotient Module]
	The \textbf{quotient module} of $M$ by $N$ is
	\[M/N \coloneqq \{m+N \mid m\in M\}\]
\end{dfn}
\begin{prop}[title =Quotient Module is \texorpdfstring{$R$}{R}-module]
	Quotient modules are $R$-modules
\end{prop}
\begin{proof}
	Define addition of cosets as 
	\[(m+N) + (m'+N) \coloneqq (m+m')+N\]
	We will write $\obar{m}$ for $m\in M$ if $N$ is understood.\\
	\Exr Check for well-definedness
	\begin{align*}
	m+N = m_1 + N \implies m-m_1=n \in N\\
	m'+N = m'_1 +N \implies m'-m'_1 = n' \in N
	\end{align*}
	Then
	\begin{align*}(m_1+N)+(m'_1+N) = (m+m'_1)+N =& (m+n+m'+n')+N\\ =& (m+m')+\underbrace{(n+n')}_{\in N} + N =(m+m')+N\end{align*}
	If $r\in R$ and $m+N \in M/N$. The $R$-action is then defined $r\rdot (m+N) \coloneqq (rm)+N$. \Exr Check it is well defined.
\end{proof}
\begin{prop}[title = Canonical quotient map]
	The natural quotient map
	\begin{align*}
	p\colon M \to& M/N \\
	m\mapsto& M+N
	\end{align*}
	is a surjective $R$-module homomorphism such that $\Ker p= N$.
\end{prop}
\begin{proof}
	Properties of an $R$-module homomorphism:
	\begin{align*}
	&p(a+b)=(a+b)+N=(a+N)+(b+N)=p(a)+p(b)\\
	&p(ra)=(ra)+N = r(a+N) = r\rdot p(a)
	\end{align*}
	Surjectivity is clear.\\
	Suppose $a\in \Ker p$, then $f(a) = a+N = 0+N$ i.e there exists $n\in N$ such that $a-0=n\in N$ and hence $a=n\in N$ so that $a\in N$.\\
	Suppose $n\in N$. Then
	\[f(n)=n+N \implies n-0\in N \implies n+N = 0+N \implies f(n) = 0+n \implies n\in \Ker p\]
\end{proof}
\begin{thm}[title = Module Isopmorphism Theorems]
	\begin{enumerate}[label=(\arabic*)]
		\item  Let $M,N$ be $R$-modules and $f\in \Hom_R(M,n)$. Then $\Ker f \subset M$ is a submodule and
		\[M/\Ker f \cong \Img f\]
		\item Let $A,B\subset M$ be submodules, then
		\[(A+B)/B \cong A/A\cap B\]
		\item Let $A\subset B\subset M$ be submodules, then
		\[(M/A)/(B/A) \cong M/B\]
		\item There is a bijection of sets
		\[\{ \text{subrings of } M \text{ containing } N \} \longleftrightarrow \{ \text{subrings of } M/N \}\]
	\end{enumerate}
	\center
	\begin{tikzcd}[column sep=0em,row sep=1em]
	&M\arrow[dl,dash]\arrow[dr,dash] &&&&& M/N\arrow[dl,dash]\arrow[dr,dash]\\
	A\arrow[dr,dash] && B\arrow[dl,dash] \arrow[rrr, leftrightarrow] % This is the new bit
	&\phantom{X}\arrow[r,dash]&\phantom{Y}& A/N\arrow[dr,dash] && B/N\arrow[dl,dash]\\
	&C\arrow[d,dash] &&&&& C/N\arrow[d,dash]\\
	&N &&&&& 0
	\end{tikzcd}   
\end{thm}
\begin{prop}[title = \texorpdfstring{$\Hom_R(M,N)$}{Hom} is an \texorpdfstring{$R$}{R}-module]
	Suppose $M,N$ are $R$-modules, then $\Hom_R(M,N)$ is itself an $R$-module
\end{prop}
\begin{proof}
	Define addition for $f,g\in \Hom_R(M,N)$ as
	\[(f+g)(m)\coloneqq f(m)+g(m)\]
	\Exr
	\begin{align*}
	0\colon M&\to N\\
	m&\mapsto 0
	\end{align*}
	is the additive identity and
	\begin{align*}
	-f\colon M&\to N\\
	m&\mapsto -f(m)
	\end{align*}
	is the additive inverse. Hence, $\Hom_R(M,N)$ is an abelian group with $+$.\\
	Then the $R$-action for $r\in R$ and $f\in \Hom_R(M,N)$ is
	\begin{align*}
	(r\rdot f)\colon M \to& N\\
	m\mapsto& r\rdot f(m)
	\end{align*}
	\Exr $\Hom_R(M,N)$ satisfies all the $R$-module action properties.
\end{proof}
\Note These operations are the same operations we learned for functions (even linear transformations). Let $f\colon \R^n \to \R^m$ be a linear transformation, then
\begin{align*}
&\begin{pmatrix}
a&b\\
c&d
\end{pmatrix} + \begin{pmatrix}
e&f\\
g&h
\end{pmatrix} = \begin{pmatrix}
a+e&b+f\\
c+g&d+h
\end{pmatrix}\\
&\alpha \begin{pmatrix}
a&b\\
c&d
\end{pmatrix} = \begin{pmatrix}
\alpha a&\alpha b\\
\alpha c&\alpha d
\end{pmatrix}
\end{align*}
\begin{prop}
	If $f\in \Hom_R(M,N), g \in \Hom_R(N,L)$ then $g\circ f \colon M\to L$ and $g\circ f \in \Hom_R(M,L)$
\end{prop}
\begin{proof}
	Check homomorphism properties
	\begin{align*}
	&g\circ f(x+y) = g(f(x+y)) = g(f(x)+f(y)) = g(f(x)) +g(f(y)) =g\circ f(x) + g\circ f(y)\\
	&g\circ f(ax) = g(f(ax)) = g(af(x)) = ag(f(x))=a\rdot g\circ f(x) \qedhere
	\end{align*}
\end{proof}
In particular, if $M=N=L$, then $f,g \in \Hom_R(M,M)$ and $g\circ f \in \Hom_R(M,M)$
\begin{crl}
	$\Hom_R(M,M)$ is a ring with $1$. In this ring, addition are $f+g$ and multiplication as $f\circ g$
\end{crl}
\begin{proof}
	We know $(\Hom_R(M,M),+)$ is an abelian group.\\
	We must check that composition is 
	\begin{enumerate}
		\item associative
		\[[(f\circ g) \circ h](x) = (f\circ g)[h(x)] = f[g(h(x))] = f[(g\circ h)(x)] = [f\circ (g\circ h)](x)\]
		\item distributes over addition
		\[[f\circ (g+h)](x) = f[(g+h)(x)] = f[g(x)+h(x)] = f(g(x)) +f(h(x)) = (f\circ g)(x) + (f\circ h)(x)\]
		\item has an identity.
		The identity map is
		\begin{align*}
		\Id \colon M&\to M \\
		m&\mapsto m
		\end{align*}
	\end{enumerate}
\end{proof}
\begin{dfn}[title = Endomorphisms and Endomorphism Ring]
	The ring $\Hom_R(M,M)$ is called the \textbf{endomorphism ring} of $M$. We sometimes denote it by $\End_R(M)$.\\
	The elements of $\End_R(M)$ are \textbf{endomorphisms}
\end{dfn}
\begin{example}
	If $M$ is any $R$-module, $a\in R$, $R$ commutative, then 
	\begin{align*}
	a\rdot \Id \colon M &\to M \\
	m&\mapsto a\rdot m
	\end{align*}
	is an endomorphism.\\
	Check
	\begin{align*}
	&(a\rdot \Id)(m+n)\coloneqq a\rdot (m+n) = a\rdot m + a\rdot n = (a\rdot \Id)(m) + (a\rdot \Id)(n)\\
	&(a\rdot \Id)(r\rdot m) \coloneqq a(r\rdot m) = (a\rdot )\rdot m = (r\rdot a)\rdot m = r\rdot (a\rdot m) = r\rdot (a\rdot \Id)(m)
	\end{align*}
\end{example}
We get a map
\begin{align*}
f\colon R&\to \End_R(M)\\
r&\mapsto r\rdot \Id
\end{align*}
\begin{claim}
	This map is a ring homomorphism.
\end{claim}
\begin{proof}
	Homomorphism properties
	\begin{align*}
	&f(r+s)\coloneqq (r+s)\rdot \Id = r\rdot \Id + s\rdot \Id = f(r)+f(s)\\
	&f(r\rdot s) = (r\rdot s)\rdot \Id = (r\rdot \Id) \rdot (s\rdot \Id) = f(r)\rdot f(s)\\
	&[(r\rdot s)\rdot \Id](m)=(r\rdot s)\rdot m = r\rdot (s\rdot m)= r\rdot (s\rdot \Id)(m) = (r\rdot \Id)\rdot (s\rdot \Id)(m)
	\end{align*}
\end{proof}
\textbf{\textcolor{BrickRed}{\underline{Warning}}}: This map is not always injective
\begin{example}
	$\modZ{4}$ is a $\Z$-module
	\begin{align*}
	f\colon \Z&\to \End(\modZ{4}) \\
	4&\mapsto 4\rdot \Id
	\end{align*}
	Then
	\[4\rdot \Id(\obar{a}) = 4\rdot \obar{a} = \obar{4a} = \obar{0} \implies 4 \in \Ker f\]
	Hence $f$ is not injective.
\end{example}
\end{document}